---
title: "STAC67A1Q1Q6"
author: "Yulun Wu"
date: "29/01/2022"
output: pdf_document
number_sections: true
---

# Q1
## (a)
```{r}
set.seed(1004912785)
Y = rnorm(20, mean=2, sd=5)
Y_bar = mean(Y)
Y_bar
sd_Y_bar = sqrt(sum((Y-Y_bar)^2)/20)
sd_Y_bar
StandardErr = sd_Y_bar/(sqrt(19)) # here is SE of mean
StandardErr
```

## (b)
```{r}
model = lm(Y ~ 1)
summary(model)
```
## These provide the same mean as (a) because E(Y) = $\beta_0$.

## (c)
```{r}
Y = replicate(10000,rnorm(20, mean=2, sd=5))
Y_bar = rowMeans(Y)
Y_bar
Y_bar_hist = hist(Y_bar)
Y_bar_hist
lines(density(Y_bar))
```
# Q6
## (a)
```{r}
Vote = read.table("vote.txt", header=T)
library(ggplot2)
ggplot(data=Vote,aes(growth,vote)) + geom_point() + stat_smooth(method="lm")
```
## (b)
```{r}
fit = lm(vote~growth, data=Vote)
summary(fit)
```
## $\hat{\beta_0}$ = 46.2476
## $\hat{\beta_1}$ = 3.0605

## (c)

```{r}
46.2476 + 0.1*3.0605
predict(fit,data.frame(growth = 0.1))
```
## The incumbent party's vote in 2008 is predicted to be 46.55365 by hand and 46.5537 by R, based on this result, "Obama" will win the election.

## (d)
$H_0:\beta_1=0$ vs $H_a:\beta_1>0$
$t=\hat{\beta_1}/SE(\hat{\beta_1})=3.0605/0.6963=`r round(3.0605/0.6963,2)`\newline$
$P-value=P(t>4.40) = `r 1-pt(4.40,14)`\newline$
So, there is an extreamly strong evidence that there is a signficant positive association between incumbent party's vote share and economical growth.

## (e)
```{r}
3.0605 + c(-1,1)*qt(0.975,14)*0.6963
```
95% CI by hand = (1.567085,4.553915)

95% CI by R:
```{r}
confint(fit)
```

## (f)
```{r}
1-pt(1/0.6963,14)
pt(-1/0.6963,14)
```
The probability that $P(|\hat{\beta_1}-\beta_1|>1)$ is 0.08646192.
