---
title: "STAC67A2"
author: "Yulun Wu"
date: "24/02/2022"
output: pdf_document
number_sections: true
---

# Q1
## (a)
```{r}
set.seed(1004912785)
Vote = read.table("vote-1.txt", header=T)
X = Vote$growth
error = rnorm(16, 0, sd=3.9)
Y  = 46.3 +4*X +error
df = data.frame(Y,X)
fit = lm(Y~X, data=df)
beta = fit$coefficients
beta 
Y0 = 47.516039 + 0.1*3.731923
Y0
predict(fit,data.frame(X = 0.1))
Sxx = sum((X-mean(X))^2)
Y0 + c(-1,1)*qnorm(0.975)*sqrt(1/16 + (0.1-mean(X))^2/Sxx)*3.9
predict(fit, newdata = data.frame(X = 0.1), interval = "confidence")

```
$\hat{\beta_0}$ = 47.516039
$\hat{\beta_1}$ = 3.731923
E(Y|$X_0$=0.1) is 47.88923 by hand and 47.88923 by R built-in function.
The 95% CI for E(Y|$X_0$=0.1) is (44.70896, 51.06950) by hand and (43.93035, 51.84812) by R built-in function.

## (b)
```{r}
beta0 = rep(0, 10000)
beta1 = rep(0, 10000)

set.seed(1004912785)

for(i in 1:10000){
error = rnorm(16, 0, sd=3.9)
Y  = 46.3 +4*X +error
df = data.frame(Y,X)
fit = lm(Y~X, data=df)
fit = lm(Y~X)
beta = fit$coefficients
beta0[i] = beta[1]
beta1[i] = beta[2]
}

par(mfrow=c(1, 2))
hist(beta0,freq = FALSE) # overlay hist with line
lines(density(beta0))
hist(beta1,freq = FALSE) 
lines(density(beta1))
result = c(mean(beta0),sd(beta0), mean(beta1), sd(beta1))
names(result) = c("Mean of Beta0", "SD of Beta0", "Mean of Beta1", "SD of Beta1")
result

Sxx = sum((X-mean(X))^2)
sd.beta0 = sqrt((1/16 + mean(X)^2/Sxx)*3.9^2)
sd.beta1 = sqrt(3.9^2/Sxx)
True.value = c(46.3, sd.beta0, 4, sd.beta1)
data.frame(True.value, result)
```
The estimated value is really closed to the theoretical value for both $\beta_0$ and $\beta_1$.

##(c)
``` {r}
#Q1(c)
set.seed(1004912785)
count = 0
for(i in 1:10000){

Vote = read.table("vote-1.txt", header=T)
X = Vote$growth
error = rnorm(16, 0, sd=3.9)
Y  = 46.3 +4*X +error
df = data.frame(Y,X)
fit = lm(Y~X, data=df)

beta = fit$coefficients
Y0 = 47.516039 + 0.1*3.731923

Sxx = sum((X-mean(X))^2)

N = predict(fit, newdata = data.frame(X = 0.1), interval = "confidence")

L = N[2]
U = N[3]
if (L<=47.88923 && 47.88923 <= U )
{
  count = count+1
}
  
  
}
print('this is the proportion ')
print(count/10000)
```
# Q2
``` {r}
x <- c(0:9)
y <- c(98, 135, 162, 178, 221, 232, 283, 300, 374, 395)
k2 = prod(y)^(1/10)
box_cox = function(lambda) {
  if (lambda == 0) {
    W = k2*log(y)
  }
  else {
    k1 = 1/(lambda*k2^(lambda-1))
    W = k1*((y^lambda)-1)
  }
  fit = lm(W~x)
  SSE = sum((fit$residuals)^2)
  SSE
  return(SSE)
}

lambda = optimize(box_cox,interval=c(-2,2))
lambda
library(MASS)
fit = lm(y~x)
result = boxcox(fit)
bilambda = result$x[which.max(result$y)]
bilambda
```
The $\lambda$ given by R built-in function is 0.5050505, and the $\lambda$ computed by function writen by me is 0.5159801.

# Q3
## (a)
```{r}
kidiq = read.csv("kidiq.csv", header=T)
fit = lm(kid.score~mom.iq,data=kidiq)
summary(fit)
n = dim(kidiq)[1]
beta = as.numeric(fit$coefficients)
Y0 = beta[1] + beta[2]*110
Sxx = sum((kidiq$mom.iq-mean(kidiq$mom.iq))^2)
SEY0 = sqrt(anova(fit)$`Mean Sq`[2]*(1/n + (110-mean(kidiq$mom.iq))^2/Sxx))
Y0 + c(-1, 1)*qt(0.975, n-2)*SEY0

predict(fit, newdata=data.frame(mom.iq=110), interval="confidence", level=0.95)

```
The 95% CI for or the mean test scores of all kids with their mother’s IQ score = 110 is (90.82506, 94.96890) by hand and (90.82506, 94.9689) by R built-in function.
## (b)
```{r}
Spred = sqrt(anova(fit)$`Mean Sq`[2]*(1+ 1/n + (110-mean(kidiq$mom.iq))^2/Sxx))
Y0 + c(-1, 1)*qt(0.995, n-2)*Spred

predict(fit, newdata=data.frame(mom.iq=110), interval="predict", level=0.99)

```
The 99% PI for  new kid’s test score when his or her mother’s IQ score = 110 is (45.55918, 140.23478) by hand and (45.55918, 140.2348) by R built-in function.
## (c)
```{r}
plot(fit$fitted.values, fit$residuals, pch=20, cex=1.5, col="blue")
abline(c(0,0))
```
No assumption violated, residual vs fitted value graph looks like random scatter.
## (d)
```{r}
qqnorm(fit$residuals)
qqline(fit$residuals)
shapiro.test(fit$residuals)
```
The normal Q-Q plot line is close to y=x line, but p-value=0.00217 < $\alpha$ = 0.05, so we reject the null hypotheses that that sample is from normal population. Sample is not from normal population.
## (e)
```{r}
library(lmtest)
bptest(fit)
```
Since p-value is 0.01398, p-value < $\alpha$ = 0.05, so we reject the null hypotheses that the residuals are distributed with equal variance. The residuals are not distributed with equal variance

## (f)
```{r}
result = boxcox(fit)
bilambda = result$x[which.max(result$y)]
bilambda
k2 = prod(kidiq$kid.score^(1/n))
k1 = 1/(bilambda*k2^(bilambda-1))
Y_star=k1*((kidiq$kid.score^bilambda)-1)
fit2 = lm(Y_star~mom.iq,data=kidiq)

qqnorm(fit2$residuals)
qqline(fit2$residuals)
shapiro.test(fit2$residuals)
bptest(fit2)
```
Since p-value from Shapiro-Wilk normality test is 0.1556>$\alpha$ = 0.05, so we fail to reject the null hypotheses that the sample is from normal population. Since p-value for Breusch-Pagan test is 0.3859, p-value=0.3859> $\alpha$ = 0.05, so we fail to reject the null hypotheses that the residuals are distributed with equal variance.

# Q4
## (a)
```{r}
x = c(0, 1, 2, 4, 6, 15)
Y = matrix(c(2.3, 4.2, 5.2, 5.9, 6.3, 7.2),nrow = 6)
X = matrix(rep(1,6),nrow = 6)
X = cbind(X,x)
X
Y
```
##(b)
```{r}
Beta_hat = solve(t(X)%*%X)%*%t(X)%*%Y
Beta_hat
Y_hat = X%*%Beta_hat
Y_hat
residual = Y-Y_hat
residual
```
##(c)
```{r}
Var_hat_beta_hat = (sum(residual^2)/4)*solve(t(X)%*%X)
Var_hat_beta_hat
```
##(d)
```{r}
H = X%*%solve(t(X)%*%X)%*%t(X)
H
library(psych)
tr(H)
```
$\sum_{i=1}^{6} h_{ii} = 2$  which is the size of $\hat{\beta}$ vector.

##(e)
```{r}
Var_hat_residual = (sum(residual^2)/4)*(diag(6)-H)
Var_hat_residual
plot(dpois(x=1:5,lambda=1))
1-ppois(3,lambda=1)
1-ppois(4,lambda=1)
1-ppois(2,lambda=1)
1-ppois(1,lambda=1)
(0.05-0.01898816)/(0.0803014-0.01898816)
1-ppois(2,lambda=1)-(1-ppois(3,lambda=1))
1-ppois(3,lambda=0.5)
(1-ppois(2,lambda=0.5))-(1-ppois(3,lambda=0.5))
0.001751623+0.01263606
1-0.01438768
```
                                                                                                                                                                                                                                                                                                                                                                            



